Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-08-04 manuscript_res.tex
add some score plots

The proposed model for ILI for any location is given in (\ref{eq:ili_model}). 
Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, 
where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given 
season. The ILI is a proportion, so the Beta random variable is a natural 
selection for modeling. Under the parameterization in of the Beta distribution 
used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the 
variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a 
scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the 
trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the 
systematic patterns which $f_{\theta_s}(w)$ does not capture. 
\begin{equation}
\begin{aligned}
    \label{eq:ili_model}
    % \begin{flalign*}
        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
    % \end{flalign*}
\end{aligned}
\end{equation}
In Osthus et al. \cite[]{osthus2019dynamic} 
$f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious 
compartment of the SIR model from (\ref{eq:sir_diff}) in section 
\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} 
$f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section 
\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
over seasons.

@@

The term $\gamma_w$, where $w$ is the season week, is included in 
(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
function. Modeling discrepancy has been used in uncertainty analysis of 
simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,
arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to 
overfitting, particularly in forecasting scenarios, and may also lead to 
identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the 
selection of prior distributions \cite[]{osthus2019dynamic,
brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
and 2016 flu seasons where their model outperformed all others in the CDC flu 
forecasting challenge \cite[]{osthus2019dynamic}. As in Osthus's model, 
$\gamma_w$ is modeled as a reverse random walk, as shown in (\ref{eq:rw}). 
\begin{equation}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} \overset{ind}{\sim} N(\gamma_{t+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W})
\end{equation}
The idea for using the reverse random walk is that there are several previous 
seasons of ILI data, and assuming the random walk captures systematic behavior, 
fitting it hierarchically over seasons can assist in predicting future behavior 
in the current season. 
Reverse random walks have also been used with success in election forecasting 
and other flu forecasting models \cite[]{osthus2021multiscale, 
osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
identifiability.

@@

The priors selected for the ILI data model under both the SIR and ASG models 
largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
numerical stability and/or we felt the adjusted prior made more sense for the 
problem. For model (\ref{eq:ili_model}), parameters which are common even when 
using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
function $\theta_s = (\alpha_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

@@

Additional prior constraints were made to improve parameter identifiability. 
In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
parameters. A modular Bayesian approach involves multiple steps of parameter 
fitting where some parameters may be estimated without priors via maximum 
likelihood estimation or other means. Fitting the rest of the model parameters 
involves assigning priors and conditioning on the previously fit parameters. 
This has been done in computer modeling to improve identifiability and other 
issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
not probabilistically sound if parameter inference is a priority 
\cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
We carried out the modular fit by first estimating the maximum likelihood 
estimate (MLE) for the parameter $\lambda_s$ for each season in 
(\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.

@@

Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison and have the property that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
The proper scoring rule primarily used in the FluSight competition and the 
COVID-19 Forecast Hub and which we will use herein is the 
weighted interval score (WIS)
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval as defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
The goal of the forecaster is to 
minimize the WIS.

@@

Figure \ref{fig:wis_by_week_horizon} shows 
boxplots of the WIS for the 500 simulation replicates faceted by season week 
on the y-axis and horizon on the x-axis. 
% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
In most plots the ASGD model shows the best performance, 
especially during week 20 where
forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
SIR model, though its performance does not match that of the ASG or ASGD.
Here the ASGD model regularly outperforms
the BASE and ARIMA models, but the three other nonlinear models do not. 
The ASG model
tends to outperform the SIR and SIRD models suggesting that the additional 
flexibility of the five parameters ASG function 
is often better for forecasting than that of the three parameter SIR function.
But the best performance comes with the combination of the ASG function and
the discrepancy modeling.

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.4]{Images/wis_by_week_horizon.png}
    \caption{Boxplots of weighted interval score (WIS) for the 
    four Bayesian nonlinear ILI models, a baseline model (BASE) and
    an autoregressive integrated moving average model (ARIMA) 
    over all seasons in the simulation study faceted by horizon 
    (x-axis) and season week (y-axis). Horizons include 1-4 week ahead forecasts 
    and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
    \label{fig:wis_by_week_horizon}
\end{figure}

@@

In this section we apply the forecast models to make forecasts of the 2023 
flu season weekly hospitalizations. The scoring of the forecasts is in the 
context of the FluSight competition where each competing forecast was 
submitted as a set of 23 quantiles corresponding to the given probability 
levels $(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975,\\ 0.990)$. A 
single forecast is thus comprised of 11 predictive intervals and a median.
Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
and forecasts were made at the state and national level. The first week of 
forecasting took place during the week of October 7, 2023, and the final week 
was the week of April 27, 2024 making 29 total weeks of forecasts.
 The same format was used during the 2021 and 2022 seasons and for the 
 COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
 Primary scores for evaluating each forecast were the WIS, 
 the log-weighted interval score (LWIS), and the relative weighted 
 interval score (RWIS). The focus in this section will be on the LWIS 
 (see supplementary material for RWIS based results \cite[]{wadsworth2024bas}).
 The LWIS is the same as the WIS except that it is evaluated 
over the log of quantiles and the log of the observed value.

We fit 24 forecast models for each location for all 29 weeks, and for each 
week forecast 1-4 week ahead hospitalizations. The 24 models included all 
combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
hospitalization models, and both quadratic and linear hospitalization models. 
The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
indicator function for values within the set $A$.

\begin{equation}
\begin{aligned}
    \label{eq:sir_prior}
        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
        \beta_s &\sim N^+(0.8, 0.3) \\
        \rho_s &\sim N^+(0.68, 0.08)
\end{aligned}
\end{equation}
For the ASG model, the MLE 
$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ was calculated and $\hat{\lambda}_s$ was accepted as a fixed value. The remaining estimates were used as starting values for posterior sampling. The transformation $T(\theta_s) = (\text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ was made, and a prior distribution was assigned to $T(\theta_s)$.
The prior distributions for the ILI model under the ASG function are shown 
in (\ref{eq:ili_prior}). These are slightly informative priors because for 
most parameters we have an idea what reasonable values may be. Here 
$m = (0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.2, 5, 2, 2)$
where $\text{diag}(\cdot)$ 
is the diagonal matrix for the given entries.

@@

\begin{equation}
\begin{aligned}
    \label{eq:shared_ili_priors}
        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2)
\end{aligned}
\end{equation}
Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
we selected an informative prior distribution by first estimating  
$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
calculating the MLE for $\theta_s$. 
% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
$\widehat{ILI}_{s,w}$ was then predicted such that 
$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
seasons of the differences 
$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
calculated using the \texttt{mle2} function in the \texttt{bblme} package 
\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
function.
The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).

@@

\begin{figure}[hbt!]
\centering
\begin{subfigure}
  \centering
  \includegraphics[width=.49\linewidth,height=6cm]{Images/normal_forecasts_us.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}
  \centering
  \includegraphics[width=.49\linewidth, height=6cm]{Images/normal_sq_forecasts_us.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
\end{subfigure}
\caption{Forecasts 1-4 weeks ahead for US hospitalizations during the 2023 
season for weeks 14, 20, 26, and 32. Forecasts are separated by ILI model, 
and the hospitalization models are all normally distribution. The figure 
includes hospitalization forecasts where ILI is a linear predictor (left) 
and where ILI is a quadratic predictor (right). 50\% predictive intervals 
are pink and 95\% predictive intervals are red.}
\label{fig:normal_flu_forecasts}
\end{figure}

Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
flu hospitalizations during the 2023 season under the NORM hospitalization 
model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive 
bands are the 50\% and 95\% predictive intervals.
These plots show a tendency to often underpredict hospitalizations. The 
quadratic models appear to do a better job predicting  hospitalizations at 
the season peak but a poorer job predicting after the peak, whereas the linear 
models seem to predict well or slightly overpredict after the peak. 
% From these figures, it's difficult to say whether ASG or SIR models predict better, but at least for the quadractic models the week 22 predictions are maybe slightly better than for the linear models.

Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the 
season for all 24 models of US hospitalizations. LWIS scores are grouped by 
ILI model, hospitalization model distribution, and by the linear or quadratic 
modeling. Here, the smaller the LWIS the better. The weeks around the holiday 
week 22 are highlighted by a grey band. 
In several cases, there appears to be a turning point in performance at or 
near week 22. For most ASG models, the models which include discrepancy tend 
to outperform those which do not beginning near week 22 and for the rest of 
the season.
For the SIR models, the discrepancy does not always lead to better forecasts, 
but in for the LNORM hospitalization models, including discrepancy greatly 
improves forecasts around the holiday week 22. The LNORM LWIS scores tend to 
be higher than for LST and NORM hospitalization models, but figure 
\ref{fig:lwis_by_dist_loc} and table \ref{tab:forecast_scores} suggest that 
the LNORM hospitalization model outperforms LST and NORM for most states and 
territories.

@@

    % \centering
    \includegraphics[scale = .6]{Images/lwis_ln_us_full_season.png}
    \caption{Each plot shows the log weighted interval score (LWIS) for every 
    week of the 2023 flu season with scores for models including and exluding 
    discrepancy in the ILI model. Scores are separated by hospitalization 
    distribution family and by ILI as a linear or quadratic predictor. Scores 
    for models with an ASG ILI model are above while those with an SIR model 
    are below. The lower the LWIS the better the forecast.}
    \label{fig:us_lwis}
\end{figure}

>>>>

The proposed model for ILI for any location is given in (\ref{eq:ili_model}). 
Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, 
where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given 
season. The ILI is a proportion, so the Beta random variable is a natural 
selection for modeling. Under the parameterization in of the Beta distribution 
used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the 
variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a 
scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the 
trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the 
systematic patterns which $f_{\theta_s}(w)$ does not capture. 
\begin{equation}
\begin{aligned}
    \label{eq:ili_model}
    % \begin{flalign*}
        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w + \upsilon_{s,w}
    % \end{flalign*}
\end{aligned}
\end{equation}
In Osthus et al. \cite[]{osthus2019dynamic} 
$f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious 
compartment of the SIR model from (\ref{eq:sir_diff}) in section 
\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} 
$f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section 
\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
over seasons.

@@

Modeling discrepancy has been used in uncertainty analysis of 
simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,
arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to 
overfitting, particularly in forecasting scenarios, and may also lead to 
identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the 
selection of prior distributions \cite[]{osthus2019dynamic,
brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
and 2016 flu seasons where their model outperformed all others in the CDC flu 
forecasting challenge \cite[]{osthus2019dynamic}.
The terms $\gamma_w$ and $\upsilon_{s,w}$, where $s$ is the season and $w$ is the 
season week, are included in 
(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
function. Here $\gamma_w$ is shared across all seasons, whereas $\upsilon_{s,w}$
is specific to season $s$. In fact, to assist in identifying $\gamma_w$,
$\upsilon_{s,w}$ is only included for the current season being forecast and none of
the previous seasons.
 As in Osthus's model, 
$\gamma_w$ and $\upsilon_{s,w}$ are modeled as reverse random walks, as shown 
in (\ref{eq:rw}). 
\begin{equation}
    \begin{aligned}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{w+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W}) \\
        \upsilon_{s,w}|\upsilon_{s, w + 1} &\overset{ind}{\sim} N(\upsilon_{s, w+1},\sigma^2_{\upsilon})
    \end{aligned}
\end{equation}

The idea for using the reverse random walk is that there are several previous 
seasons of ILI data, and assuming the random walk captures systematic behavior, 
fitting it hierarchically over seasons can assist in predicting future behavior 
in the current season. 
Reverse random walks have also been used with success in election forecasting 
and other flu forecasting models \cite[]{osthus2021multiscale, 
osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
identifiability. An additional constraint is $\upsilon_{s, W} = 0$.

@@

The priors selected for the ILI data model under both the SIR and ASG models 
largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
numerical stability and/or we felt the adjusted prior made more sense for the 
problem. For model (\ref{eq:ili_model}), parameters which are common even when 
using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
function $\theta_s = (\lambda_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

@@

Additional prior constraints were made to improve parameter identifiability. 
In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. 
% To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
% in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
% parameters. A modular Bayesian approach involves multiple steps of parameter 
% fitting where some parameters may be estimated without priors via maximum 
% likelihood estimation or other means. Fitting the rest of the model parameters 
% involves assigning priors and conditioning on the previously fit parameters. 
% This has been done in computer modeling to improve identifiability and other 
% issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
% not probabilistically sound if parameter inference is a priority 
% \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
% We carried out the modular fit by first estimating the maximum likelihood 
% estimate (MLE) for the parameter $\lambda_s$ for each season in 
% (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.

@@

Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison and have the property that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
The proper scoring rules primarily used in the FluSight competition and the 
COVID-19 Forecast Hub and which we will use herein are the 
weighted interval score (WIS) and the log-weighted interval score (LWIS).
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval as defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
The LWIS is the same as the WIS except that it is evaluated 
over the log of quantiles and the log of the observed value.

@@

The forecasts used in this section and in the subsequent section follow the 
FluSight convention of being quantiles corresponding to the 23 given 
probabilities $\\(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$
making each forecast a set of 11 predictive intervals and a predictive median.
Figure \ref{fig:wis_by_week_horizon} shows 
boxplots of the WIS for the 500 simulation replicates faceted by week 
on the y-axis and horizon on the x-axis. 
% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
In most plots the ASGD model shows the best performance, 
especially during week 20 where
forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
SIR model, though its performance does not match that of the ASG or ASGD.
Here the ASGD model regularly outperforms
the BASE and ARIMA models. The three other nonlinear models do not perform as
well.
The ASG model
tends to outperform the SIR and SIRD models suggesting that the additional 
flexibility of the five parameters ASG function 
is often better for forecasting than that of the three parameter SIR function.
The model with the best performance is the ASGD.The ASGD model shows 
especially good forecast skill at week 20 around where we would expect the 
discrepancy to be most useful.

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.57]{Images/wis_by_week_horizon.png}
    \caption{Boxplots of weighted interval score (WIS) for the 
    four Bayesian nonlinear ILI models, a baseline model (BASE) and
    an autoregressive integrated moving average model (ARIMA) 
    over all seasons in the simulation study faceted by horizon 
    (x-axis) and season week (y-axis). Horizons include 1-4 week ahead forecasts 
    and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
    \label{fig:wis_by_week_horizon}
\end{figure}

\begin{table}[hbt!]
\centering
\begin{tabular}{lrrrr}
Model & WIS & MAE & 50\%Coverage  & 95\% Coverage \\ 
  \hline
ASGD & 0.09 & 0.15 & 42 & 91 \\ 
  ASG & 0.13 & 0.18 & 33 & 71 \\ 
  ARIMA & 0.13 & 0.18 & 35 & 68 \\ 
  BASE & 0.15 & 0.20 & 33 & 65 \\ 
  SIRD & 0.15 & 0.23 & 33 & 78 \\ 
  SIR & 0.21 & 0.29 & 20 & 51 \\ 
\end{tabular}
\end{table}

@@

In this section we apply the forecast models to make forecasts for the 2023 
flu season weekly hospitalizations and compare the results with those of 
forecasts submitted to FluSight. The scoring of the forecasts is in the 
context of the FluSight competition where each competing forecast was 
submitted as a set of 23 quantiles corresponding to give probabilities.
Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
and forecasts were made at the state and national level. The first week of 
forecasting took place during the week of October 7, 2023, and the final week 
was the week of May 4, 2024 making 30 total weeks of forecasts.
 The same format was used during the 2021 and 2022 seasons and for the 
 COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
 Primary scores for evaluating each forecast were the WIS and LWIS

The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
indicator function for values within the set $A$.

\begin{equation}
\begin{aligned}
    \label{eq:sir_prior}
        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
        \beta_s &\sim N^+(0.8, 0.3) \\
        \rho_s &\sim N^+(0.68, 0.08)
\end{aligned}
\end{equation}
For the ASG model, the MLE 
$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
These estimates were used as starting values for posterior sampling. 
The transformation 
$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
was made, and a prior distribution was assigned to $T(\theta_s)$.
The prior distributions for the ILI model under the ASG function are shown 
in (\ref{eq:ili_prior}). These are slightly informative priors because for 
most parameters we have an idea what reasonable values may be. Here 
$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
where $\text{diag}(\cdot)$ 
is the diagonal matrix for the given entries.

@@

\begin{equation}
\begin{aligned}
    \label{eq:shared_ili_priors}
        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
\end{aligned}
\end{equation}
Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
we selected an informative prior distribution by first estimating  
$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
calculating the MLE for $\theta_s$. 
% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
$\widehat{ILI}_{s,w}$ was then predicted such that 
$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
seasons of the differences 
$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
calculated using the \texttt{mle2} function in the \texttt{bblme} package 
\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
function.
The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).

@@

We fit 24 forecast models for each location for all 30 weeks, and for each 
week forecast 1-4 week ahead hospitalizations. The 24 models included all 
combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
hospitalization models, and both quadratic and linear hospitalization models.
Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
flu hospitalizations during the 2023 season under the NORM hospitalization 
model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive 
bands are the 50\% and 95\% predictive intervals.
These plots show a tendency to often underpredict hospitalizations. The 
quadratic models appear to do a better job predicting  hospitalizations at 
the season peak but a poorer job predicting after the peak, whereas the linear 
models seem to predict well or slightly overpredict after the peak.

Of the 24 models, the one which produced the best forecasts was that using
the ASGD ILI model and the NORM quadratic hospitalization model. The results of
that analysis may be found in the supplementary materials ??. For the remainder 
of this section, forecast performance from the best model, which we denote
as ASG\_NORM2, is compared to that of forecasts from 20 models from which
submissions were made to FluSight in 2023. 
The models selected are those which were not ensemble models. Among the models
was the FluSight baseline. The baseline forecast had as a median the most 
recently observed 
hospitalization count, and the forecast uncertainty was based on 
differences between 
previous hospitalizations. The baseline forecasts was
similar to the baseline forecast used in
previous flu forecasting seasons and in the COVID-19 hub 
\cite[]{mathis2024evaluation, Cramer2022-hub-dataset}.

Figure \ref{fig:lwis_cover_sum} shows forecast results for 2023 flu 
hospitalization forecasts for all 21 models being compared but with results
for the ASG\_NORM2 and FluSight baseline models highlighted. 
For each week of the 2023 season, the  
LWIS averaged over location and horizon 
for each model is shown in a). Boxplots of LWIS for 1, 2, 3, and 4-week ahead
horizons for the ASG\_NORM2 and baseline models are shown in b). And 
overall empirical coverage of the 11 predictive intervals for all models is 
shown in c) with ASG\_NORM2 and the baseline model again highlighted. 
These plots show that over the season the ASG\_NORM2 model tended to 
outperform most others in terms of LWIS, and in general is superior to the 
baseline model. The coverage plot shows that the ASG\_NORM2 model is well 
calibrated at most levels, with slight overcoverage at the lower levels
up to about 80\%, and has better coverage than all other models. Similar results
for the WIS are shown in the supplementary materials ??.

@@

    % \centering
    \includegraphics[scale = .6]{Images/lwis_cover_sum.png}
    \caption{a) Averaged over location and horizon log-weighted interval score
    (LWIS) for each week during the season for the 21 models being compared. 
    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
    FluSight-baseline (red) are colored to stand out.
    b) Boxplots of all LWIS scores for 1, 2, 3, and 4-week ahead horizons 
    for the ASG\_NORM2 (blue) and FluSight-baseline (red) models.
    c) Plot of empirical coverage against theoretical coverage for 11 
    predictive intervals from the 10\% nominal level to the 98\%. The black
    line $x = y$ shows where optimal coverage would fall.
    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
    FluSight-baseline (red) are colored to stand out.}
    \label{fig:lwis_cover_sum}
\end{figure}
